COVID19 - Forecast analysis - time dependent SEIR model - Veneto

### 2020-03-15 draft_analysis_Italy.Rmd
Add home page, navigation and theme

The present analysis started from the dataset on COVID19 updated in https://github.com/pcm-dpc/COVID-19, database provided by the Italian Civil Protection.  
```{r setup}
rm(list=ls())
###import italian dataset updated 13 March 2020
dat_csv<-read.csv("https://raw.githubusercontent.com/pcm-dpc/COVID-19/master/dati-andamento-nazionale/dpc-covid19-ita-andamento-nazionale.csv",header=T)
days<-dim(dat_csv)[1]
dat_csv$t<-1:days
#database contains  the following variables
names(dat_csv)
# The total number of epidemic day is
days
plot(as.Date(substr(dat_csv$data,1,10)),dat_csv$totale_attualmente_positivi,ylab=" Infected",xlab="Date",type="l")
points(as.Date(substr(dat_csv$data,1,10)),dat_csv$totale_attualmente_positivi)
```

@@

We calculate several R0 values, each one based on a different number of days before the last day, in order to assess if the R0 trend is decreasing (how is expected to be).

```{r R0 trend}
#calculate r0 based in the last XX observation 
#vector for beta and standard deviation
beta_vec<-NULL
sd_vec<-NULL
#for cycle for all days to only the last 5.
for (i in 1:(days-4)){
fit <- lm(log(totale_attualmente_positivi)~t,data=dat_csv[i:days,])
beta_vec<-c(beta_vec,coef(fit)[2])
sd_vec<-c(sd_vec,coef(summary(fit))[2,2])
}

label <-days:5

@@

df <- data.frame(label, mean, lower, upper)[order(label),]
df$label <- factor(df$label, levels=rev(df$label))

library(ggplot2)
fp <- ggplot(data=df, aes(x=label, y=mean, ymin=lower, ymax=upper)) +
  geom_pointrange() + 
  geom_hline(yintercept=1, lty=2) +  # add a dotted line at x=1 after flip
  xlab("Last days used for the calculation") + ylab("R0 Mean (95% CI)") +
  theme_bw() 
print(fp)

@@

The R0 shows a decreasing trend in the last period. We use the estimated trend between R0 and time to calculate the future R0 value for the next 14 days. 
We predict beta (and R0) for the next 14 days assuming a Gamma distribution for the beta (the slope) forcing its value to be greater than 0. The trend was not monotonic, we use a simple splines to increase the fitting of the model to the data.
```{r R0 forecast}
library(splines)
time<-1:(days-4)
# splines with only 3 basis, weighted for standard deviation of estimates.
beta.model<-glm(beta_vec~bs(time,3),weights = 1/sd_vec,family=Gamma)
forecast=14
# add 'fit', 'lwr', and 'upr' columns to dataframe (generated by predict)
pre<-predict(beta.model,type='response',newdata=data.frame(time=1:(days+forecast)),se.fit=TRUE)
date<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+forecast-1+dim(dat_csv)[1],1)
beta.predict <- data.frame(beta_vec=c(beta_vec,rep(NA,forecast+4)),time=date,fit=pre$fit,lwr=pre$fit-1*1.96*pre$se.fit,upr=pre$fit+1*1.96*pre$se.fit)

>>>>

The present analysis started from the dataset on COVID19 updated in https://github.com/pcm-dpc/COVID-19, database provided by the Italian Civil Protection.  
```{r setup}
rm(list=ls())
###import italian dataset updated 13 March 2020
  dat_csv<-read.csv("https://raw.githubusercontent.com/pcm-dpc/COVID-19/master/dati-andamento-nazionale/dpc-covid19-ita-andamento-nazionale.csv",header=T)
days<-dim(dat_csv)[1]
dat_csv$t<-1:days
#database contains  the following variables
names(dat_csv)
# The total number of epidemic day is
days
```

@@

We calculate several R0 values, each one based on a mobile window of 5 days, that can be sufficient to estimate a local trend, in order to assess if the R0 trend is decreasing (how is expected to be).

```{r R0 trend}
#calculate r0 based with a mobile window of 5 days
#vector for beta and standard deviation
beta_vec<-NULL
sd_vec<-NULL
#for cycle for R0 estimates from days-2 to days+2
for (i in 3:(days-2)){
fit <- lm(log(totale_attualmente_positivi)~t,data=dat_csv[(i-2):(i+2),])
beta_vec<-c(beta_vec,coef(fit)[2])
sd_vec<-c(sd_vec,coef(summary(fit))[2,2])
}

label<-as.Date(substr(dat_csv$data,1,10))[3:(days-2)]

@@

df <- data.frame(label, mean, lower, upper)

library(ggplot2)
fp <- ggplot(data=df, aes(x=label, y=mean, ymin=lower, ymax=upper)) +
  geom_pointrange() +
  geom_hline(yintercept=1, lty=2) +  # add a dotted line at x=1 after flip
  xlab("Date") + ylab("R0 Mean (95% CI)") +
  theme_bw() 
print(fp)

@@

The R0 shows a decreasing trend in the last period. We use the estimated trend between R0 and time to calculate the future R0 value for the next 14 days. 
We predict beta (and R0) for the next 14 days assuming a Gamma distribution for the beta (the slope) forcing its value to be greater than 0. The trend was not monotonic, we use a simple splines to increase the fitting of the model to the data.
```{r R0 forecast}
library(splines)
time<-3:(days-2)
beta.model<-glm(beta_vec~time,weights = 1/sd_vec,family=Gamma)
forecast=14
# add 'fit', 'lwr', and 'upr' columns to dataframe (generated by predict)
pre<-predict(beta.model,type='response',newdata=data.frame(time=1:(days+forecast)),se.fit=TRUE)
date<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+forecast-1+dim(dat_csv)[1],1)
beta.predict <- data.frame(beta_vec=c(beta_vec,rep(NA,forecast+4)),time=date,fit=pre$fit,lwr=pre$fit-1*1.96*pre$se.fit,upr=pre$fit+1*1.96*pre$se.fit)

### 2020-03-15 draft_analysis_Veneto.Rmd
Add home page, navigation and theme

We calculate several R0 values, each one based on a different number of days before the last day, in order to assess if the R0 trend is decreasing (how is expected to be).

```{r R0 trend}
beta_vec<-NULL
sd_vec<-NULL
for (i in 1:(days-4)){
fit <- lm(log(totale_attualmente_positivi)~t,data=dat_csv[i:days,])
beta_vec<-c(beta_vec,coef(fit)[2])
sd_vec<-c(sd_vec,coef(summary(fit))[2,2])
}

label <-days:5
mean  <- (beta_vec*14)
lower <- ((beta_vec-1.96*sd_vec)*14)
upper <- ((beta_vec+1.96*sd_vec)*14)

df <- data.frame(label, mean, lower, upper)[order(label),]
df$label <- factor(df$label, levels=rev(df$label))

library(ggplot2)
fp <- ggplot(data=df, aes(x=label, y=mean, ymin=lower, ymax=upper)) +
  geom_pointrange() + 
  geom_hline(yintercept=1, lty=2) +  # add a dotted line at x=1 after flip
  xlab("Last days used for the calculation") + ylab("R0 Mean (95% CI)") +
  theme_bw()  # use a white background
print(fp)

@@

The R0 shows a decreasing trend in the last period. We use the estimated trend between R0 and time to calculate the future R0 value for the next 14 days. 
We predict beta (and R0) for the next 14 days assuming a Gamma distribution for the beta (the slope) forcing its value to be greater than 0. The trend was not monotonic, we use a simple splines to increase the fitting of the model to the data.
```{r R0 forecast}
library(splines)
time<-1:(days-4)

beta.model<-glm(beta_vec~bs(time,3),weights = 1/sd_vec,family=Gamma)
forecast=14
# add 'fit', 'lwr', and 'upr' columns to dataframe (generated by predict)
pre<-predict(beta.model,type='response',newdata=data.frame(time=1:(days+forecast)),se.fit=TRUE)
date<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+forecast-1+dim(dat_csv)[1],1)
beta.predict <- data.frame(beta_vec=c(beta_vec,rep(NA,forecast+4)),time=date,fit=pre$fit,lwr=pre$fit-1*1.96*pre$se.fit,upr=pre$fit+1*1.96*pre$se.fit)

>>>>

We calculate several R0 values, each one based on a mobile window of 5 days, that can be sufficient to estimate a local trend, in order to assess if the R0 trend is decreasing (how is expected to be).

```{r R0 trend}
#calculate r0 based with a mobile window of 5 days
#vector for beta and standard deviation
beta_vec<-NULL
sd_vec<-NULL
#for cycle for R0 estimates from days-2 to days+2
for (i in 3:(days-2)){
fit <- lm(log(totale_attualmente_positivi)~t,data=dat_csv[(i-2):(i+2),])
beta_vec<-c(beta_vec,coef(fit)[2])
sd_vec<-c(sd_vec,coef(summary(fit))[2,2])
}

label<-as.Date(substr(dat_csv$data,1,10))[3:(days-2)]

mean  <- (beta_vec*14)
lower <- ((beta_vec-1.96*sd_vec)*14)
upper <- ((beta_vec+1.96*sd_vec)*14)
df <- data.frame(label, mean, lower, upper)

library(ggplot2)
fp <- ggplot(data=df, aes(x=label, y=mean, ymin=lower, ymax=upper)) +
  geom_pointrange() +
  geom_hline(yintercept=1, lty=2) +  # add a dotted line at x=1 after flip
  xlab("Date") + ylab("R0 Mean (95% CI)") +
  theme_bw() 
print(fp)

@@

The R0 shows a decreasing trend in the last period. We use the estimated trend between R0 and time to calculate the future R0 value for the next 14 days. 
We predict beta (and R0) for the next 14 days assuming a Gamma distribution for the beta (the slope) forcing its value to be greater than 0. The trend was not monotonic, we use a simple splines to increase the fitting of the model to the data.
```{r R0 forecast}
time<-3:(days-2)
beta.model<-glm(beta_vec~time,weights = 1/sd_vec,family=Gamma)
forecast=14
# add 'fit', 'lwr', and 'upr' columns to dataframe (generated by predict)
pre<-predict(beta.model,type='response',newdata=data.frame(time=1:(days+forecast)),se.fit=TRUE)
date<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+forecast-1+dim(dat_csv)[1],1)
beta.predict <- data.frame(beta_vec=c(beta_vec,rep(NA,forecast+4)),time=date,fit=pre$fit,lwr=pre$fit-1*1.96*pre$se.fit,upr=pre$fit+1*1.96*pre$se.fit)
